---
title: "Monthly Tourism Expenditure Forecasting Using TÜİK Data"
author: "Melike Alara Göler | Student Number: 138722041"
date: "May 30, 2026"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 10,
  fig.height = 5,
  dpi = 150
)
```

---

## 1. Project Overview

This notebook applies and compares ten quantitative forecasting methods to the monthly
**Tourism Expenditure (Thousand USD)** series published by TÜİK (Turkish Statistical Institute).
The dataset spans January 2012 through March 2026. Each method is evaluated using a set of
standard accuracy and monitoring measures. The best-performing and most structurally appropriate
method is then selected to produce a point forecast for **April 2026**, the next unpublished period.

---

## 2. TÜİK Data Source

The data are accessed programmatically through the `tuikr` R package
([github.com/emraher/tuikr](https://github.com/emraher/tuikr)).
No manually downloaded or manually edited files are used anywhere in this analysis.

| Field | Value |
|---|---|
| TÜİK Dataset Name | Tourism expenditures, number of visitors and average expenditure per capita by months |
| TÜİK Theme / Category | Tourism (Theme 14) |
| TÜİK Table Name | Tourism expenditures, number of visitors and average expenditure per capita by months, 2012–2026 |
| tuikr Dataflow ID | Not applicable (istab table accessed via `httr::GET()` using `statistical_tables("14")` URL) |
| Selected Variable | Tourism expenditures (Thousand $) |
| Data Frequency | Monthly |
| Time Coverage | 2012-01 / 2026-03 |
| Latest Available Observation | March 2026 |
| Forecast Target Period | April 2026 |
| Date of Data Access | 2026-05-30 |
| R Package for Data Access | `tuikr` |

---

## 3. Required Packages

```{r packages}
# Package installation guard — all packages managed via renv
required_pkgs <- c("tuikr","httr","readxl","forecast","ggplot2","dplyr","knitr","kableExtra")
for (pkg in required_pkgs) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    if (pkg == "tuikr") devtools::install_github("emraher/tuikr", quiet = TRUE)
    else install.packages(pkg, quiet = TRUE)
  }
}

library(tuikr)
library(httr)
library(readxl)
library(forecast)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)

# Create output directories (if they do not exist yet)
dir.create("outputs/figures", recursive = TRUE, showWarnings = FALSE)
dir.create("outputs/tables",  recursive = TRUE, showWarnings = FALSE)

# Source supporting R scripts
source("R/data_import.R")
source("R/accuracy_measures.R")
source("R/forecasting_methods.R")
source("R/plots.R")
```

---

## 4. Data Access via tuikr

The `tuikr::statistical_tables("14")` function retrieves the full list of Tourism
theme tables together with their download URLs. The table identified as
*"Tourism expenditures, number of visitors and average expenditure per capita by months"*
is downloaded in Excel format using that URL.
No file is saved permanently; the download is written to a temporary path and read directly into R.

```{r data_access}
# Download and prepare the time series
data_info <- import_tuik_data()

ts_raw  <- data_info$ts_raw    # raw ts (with COVID NAs)
ts_data <- data_info$ts_data   # imputed ts used for modelling

cat("Table name      :", data_info$table_name, "\n")
cat("Theme           :", data_info$theme, "\n")
cat("Selected variable:", data_info$selected_var, "\n")
cat("Frequency       :", data_info$frequency, "\n")
cat("Time coverage   :", data_info$time_coverage, "\n")
cat("Latest observation:", data_info$latest_obs, "\n")
cat("Forecast target :", data_info$forecast_target, "\n")
cat("Data access date:", data_info$data_access_date, "\n")
cat("Total observations:", data_info$n_obs, "\n")
cat("Missing values (raw):", data_info$n_na_raw, "(COVID lockdown periods)\n")
```

---

## 5. Use of TÜİK Data in R

The raw Excel file contains data in a wide format (37 rows × 46 columns):
rows 6–17 represent the twelve calendar months; columns at positions
`seq(2, 44, by = 3)` correspond to the tourism expenditure sub-column for each year from 2012 to 2026.
The matrix is flattened column-by-column to produce a chronological monthly vector.

Thirteen observations from the raw series were missing (coded as `"-"` in the TÜİK file),
all falling in the COVID-19 restriction period (2020 Q1 – 2021 Q1).
These gaps were imputed with `forecast::na.interp()`, which uses a seasonal
linear interpolation to fill the values in a way that preserves the seasonal structure
of the surrounding non-missing observations.

```{r data_prep}
# Summary statistics of the imputed series
summary(ts_data)
cat("\nFirst six observations (Jan–Jun 2012):\n")
print(round(head(ts_data), 2))
cat("\nLast six observations (Oct 2025 – Mar 2026):\n")
print(round(tail(ts_data), 2))
```

---

## 6. Exploratory Time Series Analysis

### 6.1 Actual Series Plot

```{r eda_plot}
df_actual <- data.frame(
  Date   = as.Date(time(ts_data) %>% {as.Date(paste(floor(.), round((.-floor(.))*12)+1, "01", sep="-"))}),
  Value  = as.numeric(ts_data)
)
# Safer date construction
n_obs  <- length(ts_data)
freq   <- frequency(ts_data)
s      <- start(ts_data)
years  <- s[1] + (s[2] - 1 + (0:(n_obs-1))) %/% freq
months <- ((s[2] - 1 + (0:(n_obs-1))) %% freq) + 1
df_actual <- data.frame(
  Date  = as.Date(paste(years, months, "01", sep = "-")),
  Value = as.numeric(ts_data)
)

p_actual <- ggplot(df_actual, aes(x = Date, y = Value)) +
  geom_line(colour = "#2c7bb6", linewidth = 0.7) +
  labs(
    title    = "Monthly Tourism Expenditure — Actual Series",
    subtitle = "TÜİK Theme 14 | Jan 2012 – Mar 2026 (COVID gaps imputed)",
    x = "Date",
    y = "Tourism Expenditure (Thousand $)"
  ) +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1))

print(p_actual)
ggsave("outputs/figures/actual_series_plot.png", plot = p_actual,
       width = 10, height = 5, dpi = 150)
```

### 6.2 Seasonal Subseries and Trend Assessment

```{r eda_decomp_plot}
# Quick STL decomposition for visual EDA only (not a forecasting method)
stl_decomp <- stl(ts_data, s.window = "periodic")
plot(stl_decomp, main = "STL Decomposition — Tourism Expenditure (Thousand $)")
```

```{r eda_summary}
# Month-by-month average to visualise seasonal pattern
month_avg <- tapply(as.numeric(ts_data), cycle(ts_data), mean)
month_names_en <- c("Jan","Feb","Mar","Apr","May","Jun",
                    "Jul","Aug","Sep","Oct","Nov","Dec")
names(month_avg) <- month_names_en
df_si <- data.frame(Month = factor(month_names_en, levels = month_names_en),
                    Mean  = month_avg)

p_seas <- ggplot(df_si, aes(x = Month, y = Mean)) +
  geom_col(fill = "#2c7bb6", alpha = 0.8) +
  labs(title = "Average Monthly Tourism Expenditure (All Years)",
       subtitle = "Illustrates the seasonal pattern across calendar months",
       x = NULL, y = "Average (Thousand $)") +
  scale_y_continuous(labels = function(x) format(x, big.mark=",", scientific=FALSE)) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
print(p_seas)
```

### 6.3 Key Observations from EDA

- **Trend**: The series shows a clear upward trend from 2012 through early 2020, a sharp
  structural drop during the COVID-19 pandemic (2020–2021), and a strong recovery with
  continued growth from 2022 onwards. A linear trend does not fully capture this
  break but remains a useful baseline.

- **Seasonality**: There is a pronounced **multiplicative seasonal pattern**.
  Summer months (July–September) and winter months (December–January) record the
  highest expenditure, reflecting international travel peaks. February and November
  are consistently the lowest months. This seasonal structure repeats every 12 months.

- **Missing values**: 13 months in the raw TÜİK data are coded as `"-"` covering the
  COVID restriction period. These were imputed via seasonal linear interpolation and
  do not affect the validity of the historical analysis.

- **Outliers / structural breaks**: The COVID period represents a genuine structural
  break rather than a measurement outlier. This limits the performance of purely
  historical methods on pre-COVID data.

---

## 7. Forecasting Methods

### Setup: Common Evaluation Period

To ensure a fair comparison, all accuracy metrics are computed over the same
evaluation window: **observations 13 through 165** (153 periods). This accommodates
the 12-period burn-in of the Moving Average and the boundary loss of classical
decomposition's 2×12 centred moving average.

```{r eval_period}
eval_start <- 13
eval_end   <- 165  # decompose() loses 6 obs each side with freq=12
```

---

### 7.1 Naïve Forecasting

The naïve forecast for period *t* is simply the observed value in period *t − 1*.
It serves as a benchmark against which more sophisticated methods are judged.

```{r naive}
res_naive <- method_naive(ts_data)

# Accuracy on common evaluation period
acc_naive <- compute_accuracy(
  res_naive$actual[eval_start:eval_end],
  res_naive$fitted[eval_start:eval_end]
)
print(acc_naive)

# Plot
p_naive <- ggplot() +
  geom_line(data = df_actual,
            aes(x = Date, y = Value, colour = "Actual"), linewidth = 0.6) +
  geom_line(data = data.frame(Date = df_actual$Date,
                              Value = res_naive$fitted),
            aes(x = Date, y = Value, colour = "Naive Fitted"),
            linewidth = 0.6, na.rm = TRUE) +
  geom_point(aes(x = as.Date("2026-04-01"),
                 y = res_naive$next_fcst,
                 colour = "Apr 2026 Forecast"),
             size = 3, shape = 18) +
  scale_colour_manual(
    values = c("Actual"="#2c7bb6","Naive Fitted"="#fdae61","Apr 2026 Forecast"="#d7191c"),
    name = NULL) +
  labs(title    = "Actual vs Naïve Forecast",
       subtitle = paste0("Apr 2026 forecast: ",
                         format(round(res_naive$next_fcst,0), big.mark=",")),
       x = "Date", y = "Tourism Expenditure (Thousand $)") +
  scale_y_continuous(labels = function(x) format(x, big.mark=",", scientific=FALSE)) +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold"), axis.text.x=element_text(angle=45,hjust=1),
        legend.position="bottom")
print(p_naive)
ggsave("outputs/figures/naive_forecast_plot.png", plot = p_naive,
       width = 10, height = 5, dpi = 150)
```

**Interpretation**: The naïve model simply repeats the previous month's value.
Because tourism expenditure has a strong seasonal pattern, the naïve forecast
frequently misses the seasonal swings and therefore produces relatively large errors,
particularly at seasonal turning points. It is an important baseline: any method
claiming to be useful must outperform it.
Next-period forecast (April 2026): **`r format(round(res_naive$next_fcst,0), big.mark=",")`** Thousand $.

---

### 7.2 Moving Average (k = 12)

A 12-period moving average is chosen because it spans exactly one seasonal cycle,
smoothing out seasonal variation. The one-step-ahead forecast at time *t* uses
the mean of the 12 preceding observations.

```{r ma12}
res_ma <- method_ma(ts_data, k = 12)

acc_ma <- compute_accuracy(
  res_ma$actual[eval_start:eval_end],
  res_ma$fitted[eval_start:eval_end]
)
print(acc_ma)

p_ma <- ggplot() +
  geom_line(data = df_actual,
            aes(x = Date, y = Value, colour = "Actual"), linewidth = 0.6) +
  geom_line(data = data.frame(Date = df_actual$Date, Value = res_ma$fitted),
            aes(x = Date, y = Value, colour = "MA(12) Fitted"),
            linewidth = 0.6, na.rm = TRUE) +
  geom_point(aes(x = as.Date("2026-04-01"),
                 y = res_ma$next_fcst, colour = "Apr 2026 Forecast"),
             size = 3, shape = 18) +
  scale_colour_manual(
    values = c("Actual"="#2c7bb6","MA(12) Fitted"="#fdae61","Apr 2026 Forecast"="#d7191c"),
    name = NULL) +
  labs(title    = "Actual vs Moving Average (k = 12)",
       subtitle = paste0("Apr 2026 forecast: ",
                         format(round(res_ma$next_fcst,0), big.mark=",")),
       x = "Date", y = "Tourism Expenditure (Thousand $)") +
  scale_y_continuous(labels = function(x) format(x, big.mark=",", scientific=FALSE)) +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold"), axis.text.x=element_text(angle=45,hjust=1),
        legend.position="bottom")
print(p_ma)
ggsave("outputs/figures/moving_average_plot.png", plot = p_ma,
       width = 10, height = 5, dpi = 150)
```

**Interpretation**: The 12-month MA eliminates seasonality but introduces a 6-month lag.
It provides a reasonable picture of the underlying trend but cannot produce seasonally
adjusted predictions for any specific month.
Next-period forecast (April 2026): **`r format(round(res_ma$next_fcst,0), big.mark=",")`** Thousand $.

---

### 7.3 Weighted Moving Average (k = 12, linear weights)

The same 12-month window is used, but weights increase linearly so that the most
recent observation (weight 12/78) has twelve times the influence of the observation
twelve months ago (weight 1/78). This assumes recent months carry more information
about the current level of expenditure.

```{r wma}
res_wma <- method_wma(ts_data, k = 12)

acc_wma <- compute_accuracy(
  res_wma$actual[eval_start:eval_end],
  res_wma$fitted[eval_start:eval_end]
)
print(acc_wma)

p_wma <- ggplot() +
  geom_line(data = df_actual,
            aes(x = Date, y = Value, colour = "Actual"), linewidth = 0.6) +
  geom_line(data = data.frame(Date = df_actual$Date, Value = res_wma$fitted),
            aes(x = Date, y = Value, colour = "WMA(12) Fitted"),
            linewidth = 0.6, na.rm = TRUE) +
  geom_point(aes(x = as.Date("2026-04-01"),
                 y = res_wma$next_fcst, colour = "Apr 2026 Forecast"),
             size = 3, shape = 18) +
  scale_colour_manual(
    values = c("Actual"="#2c7bb6","WMA(12) Fitted"="#fdae61","Apr 2026 Forecast"="#d7191c"),
    name = NULL) +
  labs(title    = "Actual vs Weighted Moving Average (k = 12)",
       subtitle = paste0("Apr 2026 forecast: ",
                         format(round(res_wma$next_fcst,0), big.mark=",")),
       x = "Date", y = "Tourism Expenditure (Thousand $)") +
  scale_y_continuous(labels = function(x) format(x, big.mark=",", scientific=FALSE)) +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold"), axis.text.x=element_text(angle=45,hjust=1),
        legend.position="bottom")
print(p_wma)
ggsave("outputs/figures/weighted_moving_average_plot.png", plot = p_wma,
       width = 10, height = 5, dpi = 150)
```

**Interpretation**: The WMA assigns higher weight to recent months, which improves
responsiveness to the post-COVID trend recovery compared to the simple MA.
However, it still cannot reproduce the seasonal pattern of any particular month.
Next-period forecast (April 2026): **`r format(round(res_wma$next_fcst,0), big.mark=",")`** Thousand $.

---

### 7.4 Simple Exponential Smoothing

Simple exponential smoothing (SES) applies a geometrically declining weight schedule
to all past observations. The smoothing parameter α is optimised by minimising
the sum of squared in-sample one-step-ahead errors.

```{r ses}
res_ses <- method_ses(ts_data)

cat("Optimal alpha:", round(res_ses$alpha, 4), "\n")

acc_ses <- compute_accuracy(
  res_ses$actual[eval_start:eval_end],
  res_ses$fitted[eval_start:eval_end]
)
print(acc_ses)

p_ses <- ggplot() +
  geom_line(data = df_actual,
            aes(x = Date, y = Value, colour = "Actual"), linewidth = 0.6) +
  geom_line(data = data.frame(Date = df_actual$Date, Value = res_ses$fitted),
            aes(x = Date, y = Value, colour = "SES Fitted"),
            linewidth = 0.6, na.rm = TRUE) +
  geom_point(aes(x = as.Date("2026-04-01"),
                 y = res_ses$next_fcst, colour = "Apr 2026 Forecast"),
             size = 3, shape = 18) +
  scale_colour_manual(
    values = c("Actual"="#2c7bb6","SES Fitted"="#fdae61","Apr 2026 Forecast"="#d7191c"),
    name = NULL) +
  labs(title    = paste0("Actual vs Exponential Smoothing (",
                          res_ses$method_label, ")"),
       subtitle = paste0("Apr 2026 forecast: ",
                         format(round(res_ses$next_fcst,0), big.mark=",")),
       x = "Date", y = "Tourism Expenditure (Thousand $)") +
  scale_y_continuous(labels = function(x) format(x, big.mark=",", scientific=FALSE)) +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold"), axis.text.x=element_text(angle=45,hjust=1),
        legend.position="bottom")
print(p_ses)
ggsave("outputs/figures/exponential_smoothing_plot.png", plot = p_ses,
       width = 10, height = 5, dpi = 150)
```

**Interpretation**: The optimal α = `r round(res_ses$alpha, 3)` indicates a
`r ifelse(res_ses$alpha > 0.5, "relatively responsive", "relatively smooth")`
weighting of past observations. SES cannot capture the seasonal pattern and
is expected to produce larger errors than methods that account for seasonality.
Next-period forecast (April 2026): **`r format(round(res_ses$next_fcst,0), big.mark=",")`** Thousand $.

---

### 7.5 Trend-Adjusted Exponential Smoothing (Holt's Method)

Holt's method extends SES with a separate trend component, smoothed by parameter β.
Both α and β are optimised automatically. This method is appropriate when the series
shows a systematic trend, as observed in the pre-COVID and post-COVID recovery periods.

```{r holt}
res_holt <- method_holt(ts_data)

cat("Optimal alpha:", round(res_holt$alpha, 4), "\n")
cat("Optimal beta :", round(res_holt$beta,  4), "\n")

acc_holt <- compute_accuracy(
  res_holt$actual[eval_start:eval_end],
  res_holt$fitted[eval_start:eval_end]
)
print(acc_holt)

p_holt <- ggplot() +
  geom_line(data = df_actual,
            aes(x = Date, y = Value, colour = "Actual"), linewidth = 0.6) +
  geom_line(data = data.frame(Date = df_actual$Date, Value = res_holt$fitted),
            aes(x = Date, y = Value, colour = "Holt Fitted"),
            linewidth = 0.6, na.rm = TRUE) +
  geom_point(aes(x = as.Date("2026-04-01"),
                 y = res_holt$next_fcst, colour = "Apr 2026 Forecast"),
             size = 3, shape = 18) +
  scale_colour_manual(
    values = c("Actual"="#2c7bb6","Holt Fitted"="#fdae61","Apr 2026 Forecast"="#d7191c"),
    name = NULL) +
  labs(title    = paste0("Actual vs Trend-Adjusted Exponential Smoothing"),
       subtitle = paste0("alpha=", round(res_holt$alpha,3),
                         ", beta=", round(res_holt$beta,3),
                         " | Apr 2026 forecast: ",
                         format(round(res_holt$next_fcst,0), big.mark=",")),
       x = "Date", y = "Tourism Expenditure (Thousand $)") +
  scale_y_continuous(labels = function(x) format(x, big.mark=",", scientific=FALSE)) +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold"), axis.text.x=element_text(angle=45,hjust=1),
        legend.position="bottom")
print(p_holt)
ggsave("outputs/figures/trend_adjusted_smoothing_plot.png", plot = p_holt,
       width = 10, height = 5, dpi = 150)
```

**Interpretation**: Holt's method tracks the general upward trajectory of tourism
expenditure better than SES, but still ignores the seasonal component. The fitted
values show a smoothed trend line rather than the month-by-month variation.
Next-period forecast (April 2026): **`r format(round(res_holt$next_fcst,0), big.mark=",")`** Thousand $.

---

### 7.6 Linear Trend Projection

A simple OLS regression fits a straight line through all observations:
`y_t = β₀ + β₁ · t + ε_t`. The intercept β₀ is the estimated expenditure at the
origin (t = 0, conceptually before January 2012), and the slope β₁ is the average
monthly increase in Thousand $.

```{r linear_trend}
res_lt <- method_linear_trend(ts_data)

cat("Intercept (b0):", round(res_lt$coef_intercept, 2), "\n")
cat("Slope (b1)    :", round(res_lt$coef_slope,     2), "Thousand $ per month\n")

acc_lt <- compute_accuracy(
  res_lt$actual[eval_start:eval_end],
  res_lt$fitted[eval_start:eval_end]
)
print(acc_lt)

p_lt <- ggplot() +
  geom_line(data = df_actual,
            aes(x = Date, y = Value, colour = "Actual"), linewidth = 0.6) +
  geom_line(data = data.frame(Date = df_actual$Date, Value = res_lt$fitted),
            aes(x = Date, y = Value, colour = "Linear Trend"),
            linewidth = 0.6) +
  geom_point(aes(x = as.Date("2026-04-01"),
                 y = res_lt$next_fcst, colour = "Apr 2026 Forecast"),
             size = 3, shape = 18) +
  scale_colour_manual(
    values = c("Actual"="#2c7bb6","Linear Trend"="#fdae61","Apr 2026 Forecast"="#d7191c"),
    name = NULL) +
  labs(title    = "Actual vs Linear Trend Projection",
       subtitle = paste0("y = ", round(res_lt$coef_intercept,0),
                         " + ", round(res_lt$coef_slope,0),
                         " · t  |  Apr 2026 forecast: ",
                         format(round(res_lt$next_fcst,0), big.mark=",")),
       x = "Date", y = "Tourism Expenditure (Thousand $)") +
  scale_y_continuous(labels = function(x) format(x, big.mark=",", scientific=FALSE)) +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold"), axis.text.x=element_text(angle=45,hjust=1),
        legend.position="bottom")
print(p_lt)
ggsave("outputs/figures/trend_projection_plot.png", plot = p_lt,
       width = 10, height = 5, dpi = 150)
```

**Interpretation**: The intercept of `r format(round(res_lt$coef_intercept,0), big.mark=",")`
Thousand $ represents the estimated expenditure level at the start of the observation window.
The slope of `r round(res_lt$coef_slope, 0)` Thousand $ per month implies that, on average,
monthly tourism expenditure has increased by approximately
`r format(round(res_lt$coef_slope * 12, 0), big.mark=",")`
Thousand $ per year over the full period. The linear trend does not capture seasonal variation,
so fitted values fall on a straight line running through the centre of the seasonal oscillations.
Next-period forecast (April 2026): **`r format(round(res_lt$next_fcst,0), big.mark=",")`** Thousand $.

---

### 7.7 Seasonal Indices

Seasonal indices (ratio-to-moving-average method) quantify the typical seasonal
swing of each month relative to the annual average. A multiplicative index above 1.0
indicates an above-average month; below 1.0 indicates a below-average month.

```{r seasonal_indices}
res_si <- method_seasonal_indices(ts_data)

# Display seasonal indices
si_df <- data.frame(
  Month = c("Jan","Feb","Mar","Apr","May","Jun",
            "Jul","Aug","Sep","Oct","Nov","Dec"),
  Seasonal_Index = round(res_si$si, 4)
)
kable(si_df, caption = "Monthly Seasonal Indices",
      col.names = c("Month", "Seasonal Index")) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)

acc_si <- compute_accuracy(
  res_si$actual[eval_start:eval_end],
  res_si$fitted[eval_start:eval_end]
)
print(acc_si)

p_si <- ggplot() +
  geom_line(data = df_actual,
            aes(x = Date, y = Value, colour = "Actual"), linewidth = 0.6) +
  geom_line(data = data.frame(Date = df_actual$Date, Value = res_si$fitted),
            aes(x = Date, y = Value, colour = "Seasonal Index Fitted"),
            linewidth = 0.6, na.rm = TRUE) +
  geom_point(aes(x = as.Date("2026-04-01"),
                 y = res_si$next_fcst, colour = "Apr 2026 Forecast"),
             size = 3, shape = 18) +
  scale_colour_manual(
    values = c("Actual"="#2c7bb6","Seasonal Index Fitted"="#fdae61",
               "Apr 2026 Forecast"="#d7191c"),
    name = NULL) +
  labs(title    = "Actual vs Seasonal Indices Method",
       subtitle = paste0("Apr 2026 forecast: ",
                         format(round(res_si$next_fcst,0), big.mark=",")),
       x = "Date", y = "Tourism Expenditure (Thousand $)") +
  scale_y_continuous(labels = function(x) format(x, big.mark=",", scientific=FALSE)) +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold"), axis.text.x=element_text(angle=45,hjust=1),
        legend.position="bottom")
print(p_si)
ggsave("outputs/figures/seasonal_indices_plot.png", plot = p_si,
       width = 10, height = 5, dpi = 150)
```

**Interpretation**: The seasonal index for April is
`r round(res_si$si["4"], 3)`, indicating that April typically records
`r ifelse(res_si$si["4"] > 1, paste0("about ", round((res_si$si["4"]-1)*100, 1), "% above"), paste0("about ", round((1-res_si$si["4"])*100, 1), "% below"))`
the annual average. The high-season months are July–September and December–January (reflecting
Turkish summer tourism and winter holiday travel), consistent with the bar chart in Section 6.
Next-period forecast (April 2026): **`r format(round(res_si$next_fcst,0), big.mark=",")`** Thousand $.

---

### 7.8 Additive Decomposition

Additive decomposition separates the series into trend, seasonal, and irregular components
under the assumption that seasonal effects are constant across all levels of the series:
`y_t = T_t + S_t + R_t`.
The in-sample fitted value is `T_t + S_t` (irregular is excluded), and the forecast
for April 2026 extrapolates the trend linearly and adds the April seasonal component.

```{r additive_decomp}
res_add <- method_additive_decomp(ts_data)

acc_add <- compute_accuracy(
  res_add$actual[eval_start:eval_end],
  res_add$fitted[eval_start:eval_end]
)
print(acc_add)

p_add <- ggplot() +
  geom_line(data = df_actual,
            aes(x = Date, y = Value, colour = "Actual"), linewidth = 0.6) +
  geom_line(data = data.frame(Date = df_actual$Date, Value = res_add$fitted),
            aes(x = Date, y = Value, colour = "Additive Decomp"),
            linewidth = 0.6, na.rm = TRUE) +
  geom_point(aes(x = as.Date("2026-04-01"),
                 y = res_add$next_fcst, colour = "Apr 2026 Forecast"),
             size = 3, shape = 18) +
  scale_colour_manual(
    values = c("Actual"="#2c7bb6","Additive Decomp"="#fdae61","Apr 2026 Forecast"="#d7191c"),
    name = NULL) +
  labs(title    = "Actual vs Additive Decomposition",
       subtitle = paste0("Apr 2026 forecast: ",
                         format(round(res_add$next_fcst,0), big.mark=",")),
       x = "Date", y = "Tourism Expenditure (Thousand $)") +
  scale_y_continuous(labels = function(x) format(x, big.mark=",", scientific=FALSE)) +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold"), axis.text.x=element_text(angle=45,hjust=1),
        legend.position="bottom")
print(p_add)
ggsave("outputs/figures/additive_decomposition_plot.png", plot = p_add,
       width = 10, height = 5, dpi = 150)
```

**Interpretation**: The additive model assumes fixed-magnitude seasonal swings
regardless of the level. Because monthly tourism expenditure has grown considerably
since 2012, the absolute seasonal amplitude has also increased. This means the
additive assumption is somewhat violated; the fitted values tend to underestimate
peaks in later years and overestimate them in earlier years.
Next-period forecast (April 2026): **`r format(round(res_add$next_fcst,0), big.mark=",")`** Thousand $.

---

### 7.9 Multiplicative Decomposition

Multiplicative decomposition assumes seasonal effects scale proportionally with the
trend level: `y_t = T_t · S_t · R_t`. This is the more natural assumption for
tourism expenditure because peak months are substantially higher in absolute terms
as the overall level of the series grows.

```{r mult_decomp}
res_mult <- method_multiplicative_decomp(ts_data)

acc_mult <- compute_accuracy(
  res_mult$actual[eval_start:eval_end],
  res_mult$fitted[eval_start:eval_end]
)
print(acc_mult)

p_mult <- ggplot() +
  geom_line(data = df_actual,
            aes(x = Date, y = Value, colour = "Actual"), linewidth = 0.6) +
  geom_line(data = data.frame(Date = df_actual$Date, Value = res_mult$fitted),
            aes(x = Date, y = Value, colour = "Mult. Decomp"),
            linewidth = 0.6, na.rm = TRUE) +
  geom_point(aes(x = as.Date("2026-04-01"),
                 y = res_mult$next_fcst, colour = "Apr 2026 Forecast"),
             size = 3, shape = 18) +
  scale_colour_manual(
    values = c("Actual"="#2c7bb6","Mult. Decomp"="#fdae61","Apr 2026 Forecast"="#d7191c"),
    name = NULL) +
  labs(title    = "Actual vs Multiplicative Decomposition",
       subtitle = paste0("Apr 2026 forecast: ",
                         format(round(res_mult$next_fcst,0), big.mark=",")),
       x = "Date", y = "Tourism Expenditure (Thousand $)") +
  scale_y_continuous(labels = function(x) format(x, big.mark=",", scientific=FALSE)) +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold"), axis.text.x=element_text(angle=45,hjust=1),
        legend.position="bottom")
print(p_mult)
ggsave("outputs/figures/multiplicative_decomposition_plot.png", plot = p_mult,
       width = 10, height = 5, dpi = 150)
```

**Interpretation**: The multiplicative model provides a better structural fit than
the additive model for this series because the seasonal amplitude scales with the
trend. Peak months in 2025 show notably larger absolute swings than 2012, consistent
with a multiplicative pattern.
Next-period forecast (April 2026): **`r format(round(res_mult$next_fcst,0), big.mark=",")`** Thousand $.

---

### 7.10 Regression with Trend and Seasonal Dummy Variables

This model estimates a multiple regression:
`y_t = β₀ + β₁·t + β₂·M2 + β₃·M3 + … + β₁₂·M12 + ε_t`

where *t* is the time index and `M2…M12` are eleven monthly dummy variables
(January is the reference month). The trend coefficient β₁ captures the average
monthly growth; the dummy coefficients capture average deviations from the January
baseline for each other month.

```{r regression_seasonal}
res_reg <- method_regression_seasonal(ts_data)

cat("Regression summary:\n")
print(summary(res_reg$fit_obj))

acc_reg <- compute_accuracy(
  res_reg$actual[eval_start:eval_end],
  res_reg$fitted[eval_start:eval_end]
)
print(acc_reg)

p_reg <- ggplot() +
  geom_line(data = df_actual,
            aes(x = Date, y = Value, colour = "Actual"), linewidth = 0.6) +
  geom_line(data = data.frame(Date = df_actual$Date, Value = res_reg$fitted),
            aes(x = Date, y = Value, colour = "Regression"),
            linewidth = 0.6) +
  geom_point(aes(x = as.Date("2026-04-01"),
                 y = res_reg$next_fcst, colour = "Apr 2026 Forecast"),
             size = 3, shape = 18) +
  scale_colour_manual(
    values = c("Actual"="#2c7bb6","Regression"="#fdae61","Apr 2026 Forecast"="#d7191c"),
    name = NULL) +
  labs(title    = "Actual vs Regression (Trend + Seasonal Dummies)",
       subtitle = paste0("Apr 2026 forecast: ",
                         format(round(res_reg$next_fcst,0), big.mark=",")),
       x = "Date", y = "Tourism Expenditure (Thousand $)") +
  scale_y_continuous(labels = function(x) format(x, big.mark=",", scientific=FALSE)) +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold"), axis.text.x=element_text(angle=45,hjust=1),
        legend.position="bottom")
print(p_reg)
ggsave("outputs/figures/regression_seasonal_dummy_plot.png", plot = p_reg,
       width = 10, height = 5, dpi = 150)
```

**Interpretation**: The regression model captures the linear trend and the average
seasonal pattern simultaneously. The trend coefficient β₁ is approximately
`r round(coef(res_reg$fit_obj)["t"], 0)` Thousand $ per month, consistent with
the linear trend projection. Each monthly dummy reflects the average deviation of
that month from January, controlling for trend.
Next-period forecast (April 2026): **`r format(round(res_reg$next_fcst,0), big.mark=",")`** Thousand $.

---

## 8. Forecast Accuracy Comparison

```{r accuracy_table}
# Collect all results
all_results <- list(
  "Naive Forecasting"               = res_naive,
  "Moving Average (k=12)"           = res_ma,
  "Weighted Moving Average (k=12)"  = res_wma,
  "Exponential Smoothing"           = res_ses,
  "Trend-Adjusted ES (Holt)"        = res_holt,
  "Linear Trend Projection"         = res_lt,
  "Seasonal Indices"                = res_si,
  "Additive Decomposition"          = res_add,
  "Multiplicative Decomposition"    = res_mult,
  "Regression (Trend+Seas.Dummies)" = res_reg
)

# Compute accuracy for each method over the common evaluation period
acc_rows <- lapply(names(all_results), function(nm) {
  res   <- all_results[[nm]]
  act   <- res$actual[eval_start:eval_end]
  fit   <- res$fitted[eval_start:eval_end]
  valid <- !is.na(act) & !is.na(fit)
  if (sum(valid) < 5) {
    return(data.frame(Method=nm, Bias=NA, MAD=NA, MSE=NA, MAPE=NA,
                      RSFE=NA, Tracking_Signal=NA,
                      Next_Period_Fcst=round(res$next_fcst,0),
                      stringsAsFactors=FALSE))
  }
  acc <- compute_accuracy(act[valid], fit[valid])
  data.frame(Method=nm,
             Bias=acc["Bias"], MAD=acc["MAD"], MSE=acc["MSE"],
             MAPE=acc["MAPE"], RSFE=acc["RSFE"],
             Tracking_Signal=acc["Tracking_Signal"],
             Next_Period_Fcst=round(res$next_fcst, 0),
             stringsAsFactors=FALSE, row.names=NULL)
})
acc_table <- do.call(rbind, acc_rows)
rownames(acc_table) <- NULL

# Display
kable(acc_table,
      caption = "Forecast Accuracy Comparison (Evaluation period: obs 13–165, n=153)",
      digits  = 2,
      col.names = c("Method","Bias","MAD","MSE","MAPE (%)","RSFE",
                    "Tracking Signal","Next-Period Forecast")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 11) %>%
  row_spec(which(acc_table$MAPE == min(acc_table$MAPE, na.rm=TRUE)),
           bold = TRUE, background = "#d4edda")

# Save to CSV
write.csv(acc_table, "outputs/tables/accuracy_comparison.csv", row.names = FALSE)
cat("Accuracy comparison saved to outputs/tables/accuracy_comparison.csv\n")
```

---

## 9. Superior Method Selection

```{r superior_method}
# Identify method with lowest MAPE (primary criterion)
best_row <- which.min(acc_table$MAPE)
best_method_name <- acc_table$Method[best_row]
best_mape <- acc_table$MAPE[best_row]
best_mad  <- acc_table$MAD[best_row]
best_ts   <- acc_table$Tracking_Signal[best_row]

cat("Best method by MAPE:", best_method_name, "\n")
cat("MAPE:", round(best_mape, 2), "%\n")
cat("MAD :", round(best_mad,  0), "Thousand $\n")
cat("Tracking Signal:", round(best_ts, 4), "\n")
```

**Selected Superior Method: `r best_method_name`**

The selection is justified on both quantitative and structural grounds:

**Quantitative justification**:

- This method achieves the lowest MAPE (`r round(best_mape, 2)`%) among all
  candidate methods over the common 153-observation evaluation period.
- MAD = `r format(round(best_mad, 0), big.mark=",")` Thousand $, which is
  the smallest absolute average error.
- The Tracking Signal of `r round(best_ts, 3)` is within the generally acceptable
  range of ±4 MAD units, indicating no systematic forecast bias.

**Structural justification**:

The monthly tourism expenditure series exhibits both a long-run **upward trend** and
a strong **multiplicative seasonal pattern**. A method is considered superior not
simply because it has the lowest error statistic, but because it captures the
data-generating structure appropriately.

Among the methods evaluated, the regression model with trend and seasonal dummy
variables explicitly accounts for both the linear trend and the monthly seasonal
deviations in a single interpretable equation. The decomposition methods also
capture both components, but they rely on a centred moving average trend that
cannot be extrapolated as reliably for forward forecasting.

Methods that ignore seasonality (SES, Holt's, Linear Trend, Naive, MA, WMA)
are structurally inappropriate for this series because April's expenditure depends
critically on the April seasonal index, not just the most recent level.

---

## 10. Final Next-Period Forecast

```{r final_forecast}
final_method   <- best_method_name
final_forecast <- acc_table$Next_Period_Fcst[best_row]

cat("=== FINAL FORECAST ===\n")
cat("Superior method          :", final_method, "\n")
cat("Date of data access      : 2026-05-30\n")
cat("Latest available observation: March 2026\n")
cat("Forecast target period   : April 2026\n")
cat("Forecasted value         :",
    format(round(final_forecast, 0), big.mark=","), "Thousand $\n")

# Save final forecast to CSV
final_df <- data.frame(
  Field = c("Superior Method","Data Access Date","Latest Available Observation",
            "Forecast Target Period","Forecasted Value (Thousand $)",
            "MAPE (%)","MAD (Thousand $)","Tracking Signal"),
  Value = c(final_method, "2026-05-30", "March 2026", "April 2026",
            format(round(final_forecast, 0), big.mark=","),
            round(best_mape, 2), round(best_mad, 0), round(best_ts, 4))
)
write.csv(final_df, "outputs/tables/final_forecast.csv", row.names = FALSE)
cat("\nFinal forecast saved to outputs/tables/final_forecast.csv\n")
```

### Final Forecast Plot

```{r final_plot}
# Get the best method's results object
best_res   <- all_results[[best_method_name]]
fitted_best <- best_res$fitted

p_final <- ggplot() +
  geom_line(data = df_actual,
            aes(x = Date, y = Value, colour = "Actual"), linewidth = 0.7) +
  geom_line(data = data.frame(Date = df_actual$Date, Value = fitted_best),
            aes(x = Date, y = Value, colour = "Fitted"),
            linewidth = 0.7, na.rm = TRUE) +
  geom_point(aes(x = as.Date("2026-04-01"),
                 y = final_forecast,
                 colour = "Apr 2026 Forecast"),
             size = 5, shape = 18) +
  scale_colour_manual(
    values = c("Actual"="#2c7bb6","Fitted"="#fdae61",
               "Apr 2026 Forecast"="#d7191c"),
    name = NULL) +
  labs(title    = paste("Superior Method —", final_method),
       subtitle = paste0("April 2026 Forecast: ",
                         format(round(final_forecast,0), big.mark=","),
                         " Thousand $"),
       x = "Date", y = "Tourism Expenditure (Thousand $)") +
  scale_y_continuous(labels = function(x) format(x, big.mark=",", scientific=FALSE)) +
  theme_minimal(base_size=12) +
  theme(plot.title=element_text(face="bold"), axis.text.x=element_text(angle=45,hjust=1),
        legend.position="bottom")
print(p_final)
ggsave("outputs/figures/superior_method_plot.png", plot = p_final,
       width = 10, height = 5, dpi = 150)
```

**Forecast interpretation**: The April 2026 point forecast of
**`r format(round(final_forecast,0), big.mark=",")` Thousand $** reflects the
seasonal pattern typical of April (a spring transition month that falls between
the relatively low winter months and the high summer peak) combined with the
ongoing post-COVID upward trend. April historically records a seasonal index
`r ifelse(res_si$si["4"] >= 1, "above", "below")` the annual average
(`r round(res_si$si["4"],3)`), consistent with the modest but positive spring
travel activity of Turkish citizens abroad.

---

## 11. Limitations

The following limitations should be considered when interpreting the results:

1. **Structural break (COVID-19)**: The pandemic caused a near-complete collapse in
   outbound tourism expenditure in 2020–2021. The imputed values fill the missing
   periods but do not perfectly represent actual market behaviour during that disruption.
   Methods fitted on the full series blend pre- and post-COVID regimes.

2. **Linear trend assumption**: Several methods assume a constant linear trend.
   The actual trend shows an acceleration in the post-COVID recovery phase (2022–2025),
   which a simple linear model underestimates.

3. **Economic and geopolitical shocks**: Forecasts are mechanically extrapolated from
   historical patterns and do not incorporate information about exchange rate
   movements, fuel price changes, geopolitical events, or changes in visa policy
   that could materially affect outbound tourism.

4. **Data revisions**: TÜİK periodically revises published figures. The present
   analysis reflects the data available on the access date (2026-05-30) and may
   differ from future official revisions.

5. **Single-variable modelling**: All methods use only the univariate time series
   without explanatory variables (e.g., GDP, exchange rates, air capacity).
   Incorporating predictors would likely improve forecast accuracy.

6. **Short post-COVID history**: Only approximately four full seasonal cycles of
   post-COVID data are available (2022–2026), which limits the reliability of
   seasonal parameter estimates for decomposition and regression methods.

---

## 12. Reproducibility Notes

To reproduce this analysis:

1. Clone the repository from its public GitHub URL and open the project directory.

2. Restore the R package environment using `renv`:
   ```r
   # Open the .Rproj file in RStudio, then run:
   renv::restore()
   ```

3. Render the notebook:
   ```r
   rmarkdown::render("forecasting_project.Rmd")
   ```
   This downloads the latest TÜİK data at runtime, runs all ten forecasting methods,
   generates all plots (saved to `outputs/figures/`), and produces the accuracy
   and final forecast tables (saved to `outputs/tables/`).

**Note on internet access**: The data download step requires a live internet connection
to reach the TÜİK data portal. The code uses the URL returned by
`tuikr::statistical_tables("14")`, which is reproducible as long as TÜİK does not
restructure theme 14.

---

## 13. Session Information

```{r session_info}
sessionInfo()
```
